A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2019-01-08 doc/large_strain.tex
clarify what's noverl in GMG section and what's used from deal.II

For this work we adopt the matrix-free level operators within the GMG and employ Chebyshev pre- and post-smoothing on each level. In general, given a heterogeneous non-linear material, the definition of transfer operators\footnote{Finite dimensional linear operators from a vector space associated with the fine-scale mesh to the coarse-scale mesh, and in the opposite direction.} (for restriction and prolongation) and level operators is not trivial. One approach is to start from an arbitrary heterogeneous material at the fine scale and employ homogenization theory \cite{Suquet1987, Hill1972,Hashin1983,Castaneda1997} to design suitable transfer operators \cite{Miehe2007}\footnote{
  We note that this approach is limited to the case where for each quadrature point within an element the fourth-order tangent stiffness tensor is the same.
}.
In this approach, the fine-scale displacement is decomposed into long-wave and short-wave contributions by an additive split, where the former is associated to the homogeneous response of a patch of elements and the latter represents fluctuations due to the heterogeneous structure of the patch.
However, for a patch of elements consisting of the same material, the transfer operators reduce to the standard geometric transfer.
Based on this observation we adopt the following approach: It is assumed that the mesh at the coarsest level can accurately describe the heterogeneous finite-strain elastic material.
In this case, the restriction and prolongation operations for patches of elements are always performed for the same material and thus admits usage of standard geometric transfer operators.
We define the level operators as follows: Recall that the fine scale tangent operator \eqref{eq:algebraic_tangent} is obtained by linearization around the current displacement field $\gz u^h$. We then restrict this displacement field to all multigrid levels $\{l\}$ and evaluate tangent operators using the matrix-free algorithms from Section \ref{sec:mf}.
Essentially, level operators correspond to the linearization around the smoothed representation of displacement field\footnote{
Note that this differs from the classical approach where the (tangent) operator $\gz A^{l+1}$ on level $l+1$ is directly related to the (tangent) operator $\gz A^{l}$ on level $l$ via $\gz A^{l+1}=\gz I^{l+1}_{l} \gz A^l \gz I^l_{l+1}$, where $\gz I^l_{l+1}$ and $\gz I^{l+1}_l$ are global prolongation (coarse-to-fine) and restriction (fine-to-coarse) operators, respectively.
}.
Finally, at the coarsest level we use a CG solver to reduce the residual by three orders of magnitude within (at most) one hundred steps.
The multigrid algorithm is a V-cycle with two pre- and post-smoothing steps.
For examples of other approaches that combine homogenization and multigrid methods, see \cite{Bayreuther2003,Fish1995} for applications using small strain elastic materials and \cite{Kaczmarczyk2010} with applications in fracture.

>>>>

For this work we adopt the matrix-free level operators within the GMG and employ Chebyshev pre- and post-smoothing on each level. In general, given a heterogeneous non-linear material, the definition of transfer operators\footnote{Finite dimensional linear operators from a vector space associated with the fine-scale mesh to the coarse-scale mesh, and in the opposite direction.} (for restriction and prolongation) and level operators is not trivial. One approach is to start from an arbitrary heterogeneous material at the fine scale and employ homogenization theory \cite{Suquet1987, Hill1972,Hashin1983,Castaneda1997} to design suitable transfer operators \cite{Miehe2007}\footnote{
  We note that this approach is limited to the case where for each quadrature point within an element the fourth-order tangent stiffness tensor is the same.
}.
In this approach, the fine-scale displacement is decomposed into long-wave and short-wave contributions by an additive split, where the former is associated to the homogeneous response of a patch of elements and the latter represents fluctuations due to the heterogeneous structure of the patch.
However, for a patch of elements consisting of the same material, the transfer operators reduce to the standard geometric transfer.
Based on this observation we adopt the following approach: It is assumed that the mesh at the coarsest level can accurately describe the heterogeneous finite-strain elastic material.
In this case, the restriction and prolongation operations for patches of elements are always performed for the same material and thus admits usage of standard geometric transfer operators.
We define the level operators as follows: Recall that the fine scale tangent operator \eqref{eq:algebraic_tangent} is obtained by linearization around the current displacement field $\gz u^h$. We then restrict this displacement field to all multigrid levels $\{l\}$ and evaluate tangent operators using the matrix-free algorithms from Section \ref{sec:mf}.
Essentially, level operators correspond to the linearization around the smoothed representation of displacement field\footnote{
Note that this differs from the classical approach where the (tangent) operator $\gz A^{l+1}$ on level $l+1$ is directly related to the (tangent) operator $\gz A^{l}$ on level $l$ via $\gz A^{l+1}=\gz I^{l+1}_{l} \gz A^l \gz I^l_{l+1}$, where $\gz I^l_{l+1}$ and $\gz I^{l+1}_l$ are global prolongation (coarse-to-fine) and restriction (fine-to-coarse) operators, respectively.
}.
Finally, at the coarsest level we use a CG solver to reduce the residual by three orders of magnitude within (at most) one hundred steps.
The multigrid algorithm is a V-cycle with two pre- and post-smoothing steps.

{\color{red}
To summarize, in this section we propose a method by which to construct level tangent operators for finite-strain elasticity to be employed within the GMG preconditioner.
The other parts of the GMG preconditioner such as Chebyshev smoother, matrix-free interlevel transfer operators and the coarse level iterative solver are provided by the \texttt{deal.II} library.
We note that at the time of writing, $p-$ version of the GMG is not available within the \texttt{deal.II} library.
}

For examples of other approaches that combine homogenization and multigrid methods, see \cite{Bayreuther2003,Fish1995} for applications using small strain elastic materials and \cite{Kaczmarczyk2010} with applications in fracture.
